Analysis methods for characterizing the climate sensitivity of annual growth and woody productivity using tree-rings

### 2020-03-17 methods/analysis_methods.Rmd
update methods

- **ring width (radial growth)** - Annual ring width from individual-level chronologies. Helene Muller-Landau argues that, from tree scaling/geometry perspective, this metric is the closest to being size-independent (*not* BAI!) (see [Francis et al. (2017)](https://onlinelibrary.wiley.com/doi/full/10.1111/geb.12604)). Empirically, this seems to be the case in closed-canopy forests.

@@

- **basal area increment (BAI)** - (Currently don't see good reason to use this.) BAI is commonly considered by dendrochronologists to be independent of DBH (or at least moresy than ring width) after trees pass a juvenile growth phase, the logic being that if a tree produces a constant amount of stem wood area each year, BAI will hold steady and ring width will decline. However, while this may hold true in some cases such as very old conifers, in most closed-canopy forests around the world, radial/ diameter growth *increases* with DBH (e.g., Muller-Landau et al. 2006, [Anderson-Teixeira et al. 2015](https://www.researchgate.net/publication/277918165_Size-related_scaling_of_tree_form_and_function_in_a_mixed-age_forest)). There are some likely differences between trees represented by tree-rings (survivors) and the community as a whole, but in [Helcoski et al. 2019](https://nph.onlinelibrary.wiley.com/doi/abs/10.1111/nph.15906) we show a positive relationship of radial growth to DBH from tree-rings. The implication--based on basic geometry--is that BAI increases even more strongly with DBH than does radial growth, and hence BAI should *not* be considered independent of DBH. Furthermore, BAI in itself is not very meaningful, as wood is added in three dimensions, and BAI only accounts for two. Thus, KAT sees no *a priori* motivation to use BAI over ring width. BAI was used by and Alexander et al. (in review) but without assumption that BAI is independent of DBH.

>>>>

- **ring width (radial growth)** - Annual ring width from individual-level chronologies. Helene Muller-Landau argues that, from tree scaling/geometry perspective, this metric is the closest to being size-independent (as opposed to BAI) (see [Francis et al. (2017)](https://onlinelibrary.wiley.com/doi/full/10.1111/geb.12604)). Empirically, this seems to be the case in closed-canopy broadleaf forests.

@@

- **basal area increment (BAI)** - (Currently don't plan to use this.) BAI is commonly considered by dendrochronologists to be independent of DBH (or at least moresy than ring width) after trees pass a juvenile growth phase, the logic being that if a tree produces a constant amount of stem wood area each year, BAI will hold steady and ring width will decline. However, while this may hold true in some cases, evidence from closed-canopy broadleaf forests around the world (ForestGEO) suggests that radial/ diameter growth *increases* with DBH (e.g., Muller-Landau et al. 2006, [Anderson-Teixeira et al. 2015](https://www.researchgate.net/publication/277918165_Size-related_scaling_of_tree_form_and_function_in_a_mixed-age_forest)). There are some likely differences between trees represented by tree-rings (survivors) and the community as a whole, but in [Helcoski et al. 2019](https://nph.onlinelibrary.wiley.com/doi/abs/10.1111/nph.15906) we show a positive relationship of radial growth to DBH from tree-rings. The implication--based on basic geometry--is that BAI increases even more strongly with DBH than does radial growth hence should not be considered independent of DBH. Furthermore, BAI in itself is not very meaningful, as wood is added in three dimensions, and BAI only accounts for two. Thus, KAT sees no *a priori* motivation to use BAI over ring width or biomass. BAI was used by and Alexander et al. (in review) but without assumption that BAI is independent of DBH.
